Towards Transparency and Open Science

### 2023-01-06 manuscript.Rmd
refine notation further

$$
\mathcal{L}_{KL}(f, g) =\int_{\mathcal{X}}f(x)\log\left(g(x)^{-1}\right) - \int_{\mathcal{X}}f(x)\log\left(f(x)^{-1}\right)
$$

@@

To allow for quantities to be induced, we must assume that our theory is parametrized, e.g., $g(x|\theta)$.
That means our theory implies a multitude of possible probability distributions that may describe reality.
Though this assumption may be lifted, we assume each implied version of the theory is equally probable a priori.
This parametrization captures the idea that some assumptions necessary for a theory to make testable statements are arbitrary.
Of those arbitrary assumptions, we want to find those that fit the reality with the least amount of information lost.
The best parametrization is achieved by:

@@

Of course, we usually do not know $\theta_*$.
That is why it is necessary to induce it from data, denoted as $\hat\theta(y)$, where $Y \sim f^n$, that is it follows the empirical distribution.
The crucial point here is to understand what happens when we can not derive $\theta$ deductively but must substitute it inductively with an estimate $\hat\theta(y)$.
Any estimated parameters $\hat{\theta}(y)$ would almost surely not be equal to $\theta_*$ (assuming $\theta$ may take an infinite number of values, i.e., is continues).
It follows that almost surely:

@@

$$
\mathbb{E}_{X \sim f}[log(g(x|\hat{\theta}(y)))] > \mathbb{E}_{X \sim f}\left[log(g(x|\theta_o))\right] 
$$

That is to say, any induced estimate will be sub-optimal.
The inference goal, however, is to compare the theory $g$ to reality $f$, not to evaluate the estimates of $\hat\theta$.
The point is to make a statement about the theory, not to make a statement about the data in combination with the theory.
If the estimate of $\hat\theta(y)$, i.e., the inductive process, is unbiased, we may form an expectation over the data $y$:

@@

We might notice that we often get an unbiased estimate for $\mathbb{E}_{Y \sim f^n}[\log(g(y|\hat{\theta}(y)))]$, e.g., in maximum likelihood estimation.
The expectation over the data together with Taylor series expansion yields:

@@

@@

$$
\mathbb{E}_Y\mathbb{E}_X[log(g(x|\hat\theta(y)))] \approx \mathbb{E}_X[log(g(x|\hat\theta(y)))] + \mathcal{C}
$$

@@

$$
\mathcal{L}(y, g(y|\theta)) = \mathbb{E}_yL(y, g(y|\theta))
$$

@@

$$
\mathbb{E}_y L(y, g(y|\hat\theta(y)))
 = \sum_{i=1}^{n}L(y_i, g(y_i|\hat\theta(y_{-i})))\frac{1}{n}
$$

>>>>

$$
\mathcal{L}_{KL}(f, g) =\int_{\mathcal{X}}f(x)\log\left(f(x)\right)\mathrm{d}x - \int_{\mathcal{X}}f(x)\log\left(g(x)\right)\mathrm{d}x
$$

@@

To allow for quantities to be induced, we must assume that our theory is parameterized, e.g., $g(x|\theta)$.
That means our theory implies a multitude of possible probability distributions that may describe reality.
Though this assumption may be lifted, we assume each implied version of the theory is equally probable a priori.
This parametrization captures the idea that some assumptions necessary for a theory to make testable statements are arbitrary.
Of those arbitrary assumptions, we want to find those that fit the reality with the least amount of information lost.
The best parametrization is achieved by:

@@

Of course, we usually do not know $\theta_*$.
That is why it is necessary to induce it from data, denoted as $\hat\theta(y)$, where $Y$ is independently sampled from $X\sim f$.
The crucial point here is to understand what happens when we can not derive $\theta$ deductively but must substitute it inductively with an estimate $\hat\theta(y)$.
Any estimated parameters $\hat{\theta}(y)$ would almost surely not be equal to $\theta_*$ (assuming $\theta$ may take an infinite number of values, i.e., is continues).
It follows that almost surely:

@@

$$
\mathbb{E}_{X \sim f}\left[\log(g(x|\theta_*))\right] >\mathbb{E}_{X \sim f}[\log(g(x|\hat{\theta}(y)))] 
$$

That is to say, any induced estimate will be sub-optimal.
The inference goal, however, is to compare the theory $g$ to reality $f$, not to evaluate the estimates of $\hat\theta$.
The point is to make a statement about the theory, not to make a statement about the data in combination with the theory.
If the estimate of $\hat\theta(y)$, i.e., the inductive process, is unbiased in the sense that it converges towards $\theta_*$, we may form an expectation over the data $Y$:

@@

We might notice that we often get an unbiased estimate for the observed distance on the data used to induce $\hat \theta$.
With slight abuse of notation, $log(g(y|\theta)) \equiv \sum_{i = 1}^n log(g(y_i|\theta))$, so that $\mathbb{E}_{y \sim f^n}log(g(y|\hat\theta(y)))$ refers to the observed likelihood.
The observed likelihood is often easy enough to obtain, e.g. in maximum likelihood estimation.

The expectation over the data together with Taylor series expansion yields:

@@

Where $J$ is the Fisher information matrix with regard to $g$, and $I$ for $f$ respectively.
For more details see of this derivation see @burnhamModelSelectionMultimodel2002, Ch. 7.2.

@@

$$
\mathbb{E}_{X \sim f}\mathbb{E}_{Y \sim f^n}[log(g(x|\hat\theta(y)))] \approx \mathbb{E}_{Y \sim f^n}[\log(g(y|\hat{\theta}(y)))] + \mathcal{C}
$$

@@

$$
\mathcal{L}(x, g(x|\theta)) = \mathbb{E}_xL(x, g(x|\theta))
$$

@@

$$
\mathbb{E}_{x \sim f} L(x, g(x|\hat\theta(y)))
 = \frac{1}{n}\sum_{i=1}^{n}L(y_i, g(y_i|\hat\theta(y_{-i})))
$$
